Proposed Solution

### 2017-04-16 OOPSLA17/secs/learner.tex
fixed rule def to be entierly consistent (i hope)

A \textit{rule}, $r$, takes the form of an implication relationship, $r = S \in \validSet \implies T \in \validSet \land p(S,T)$, between two possibly empty sets of keywords $S,T \subset \keys$ where \validSet is the set of all valid configuration files.
In keeping with vocabulary from association rule learning, we refer to $S$ and $T$ as the source and target keyword sets respectively.
A rule states that if the source set of keyword appears in a valid configuration file, the target set of keywords must be present and the given predicate must hold.
For simplicity of notation, we treat the $S,T \in \validSet$ as implicit when writing a rule.

Taking the food store example, our learned rule would be $\{bread, butter\} \implies assoc(\{bread,butter\},\{jelly\})$.
The set $\keys$ is the set of unique keys from the training set (denoted \trainingSet) and the predicate $p$ is one of the classes of configuration errors.
The task of the learning algorithm is to transform a training set to a set of rules, weighted with \textit{support} and \textit{confidence}.
The set of rules learned from training set \trainingSet constitutes a specification for a configuration file to be considered correct.

The two metrics, support and confidence, are used in association rule learning, as well as other rule based machine learning techniques~\cite{han2007frequent,langley1995applications}.
We use slightly modified definitions of these, stated below, to handle arbitrary predicates as rules.
During the learning process, each rule is assigned a support and confidence to measure the amount and quality of evidence for the rule.
%
\begin{equation*}
 \textit{support(r)} = \frac{|\{C \in \trainingSet \mid S_r \cup T_r \subseteq C\}|} {|\trainingSet|}
\end{equation*}
\begin{equation*}
 \textit{confidence(r)} = \frac{|\{C \in \trainingSet \mid p_r(S_r,T_r) \subseteq C\}| } {support(r)*|\trainingSet|}
\end{equation*}

@@

\begin{align}
Learn(\trainingSet) = \{r | \ &support(r) > t_s\ \land \nonumber \\
    & \textit{confidence(r)} > t_c\ \land \label{eq:learn}\\ 
    & S,\ T \subset \keys\} \nonumber
\end{align}

\subsection{Error Classes}
Each class of error forms a rule with a predicate $p$, and additional restrictions on the sets $S, T$.
Ordering errors require $|S|, |T| = 1$ and use the predicate $order$ to mean the keyword $S$ must come before the keyword $T$ in any configuration file.
For example, the ordering error given in Sec.~\ref{sec:motiv} is expressed:
%
\begin{equation*}
\{\texttt{innodb\_data\_home\_dir}\} \in \validSet \implies ord(\{\texttt{innoddb\_data\_home\_dir}\},\{\texttt{innodb\_data\_file\_path}\})
\end{equation*}
%
Missing keyword entry errors require $|S|, |T| = 1$ and use the predicate $missing$ to mean the keyword $T$ must appear in the same file as the keyword $S$ in any configuration file.
The type rule is a set of rules over multiple predicates, where $|S|=1, |T|=0$.
These type rules take the form $S \implies is\_(S)$ where $\_$ matches all the basic types (bool, int, size, etc).

\app also supports two types of integer correlation rules, coarse-grained and fine-grained.
Both integer correlation rules are set of rules over the predicates $\{<,=,>\}$.
Coarse grain rules require $|S|, |T| = 1$, and the predicates have the typical interpretation.
Fine-grained rules use $|S|=2,|T|=1$, and interpret the predicates such that for $k_1,\ k_2 \in S,\ k1*k2$ must have the predicate relation to $T$.
The integer correlation rules also use probabilistic types to prune the search space.
To avoid learning too many false positives, we restrict this rule to either $size*int=size$, or $int*int = int$, as shown in Fig.~\ref{fig:allrules}.
Without probabilistic typing, we would also learn rules which do not have a invalid sematinc interpretation, for example a relation between three size keywords ($k1,k2,k3::size$) of the form $k1*k2>k3$.

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
%\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=bool]
{k1 :: bool}
{isbool(k1) :: Rule}
\and\quad
\inferrule* [Right=int]
{k1 :: int}
{isint(k1) :: Rule}
\and\\
\inferrule* [Right=fine\_grain\_rule]
{k1,k2,k3 :: int}
{fine(k1,k2,k3) :: Rule}
\and\quad
\inferrule* [Right=fine\_grain\_rule]
{k1,k3 :: size \\ k2 :: int}
{fine(k1,k2,k3) :: Rule}
\end{mathpar}
\caption{An expanded set of typing judgements for valid rules}
\label{fig:allrules}
\end{figure}
}

>>>>

In traditional association rule learning, rules take the form of an implication:
\begin{align*}
r = \{S_0,...,S_{|S|}\} \in\ valid\ \implies \{T_0,...,T_{|T|}\} \in\ valid
\end{align*}
where $S$ and $T$ are source and target sets of words.
This rule states that if the set of words $S$ appear in some $valid$ (i.e. non-rule breaking) file, the words $T$ must also appear in that file.
We generalize this relation so that a rule, $r$, instead takes the following form:
%
\begin{align*}
r = [S_0,...,S_{|S|}] \in \validSet \implies \validSet \vdash p([S_0,...,S_{|S|}],[T_0,...,T_{|T|}])
\end{align*}
%
This rule relates the keywords $S_i,T_j \subseteq \keys$ for $0 \leq i \leq |S|, 0 \leq j \leq |T|$ with a predicate $p$.
Whenever the set of keywords $S$ is present in some valid configuration file, \validSet, the predicate $p$ must hold over $S$ and $T$ given the context of \validSet.
In keeping with vocabulary from association rule learning, we refer to $S_r$ and $T_r$ as the source and target keyword lists of rule $r$ respectively.
A rule states that if the source list of keyword appears in a valid configuration file, the given predicate must hold between the source and target lists.
For clarity in notation, we define all predicates to have arity 2 by separating the source and target keywords into lists of length $|S|$ and $|T|$ respectively,
   but these can be equivalently thought of as predicates of arity $|S|+|T|$ as notated in Fig.~\ref{fig:ptypes}.

Taking the food store example, our learned rule would be:
\begin{align*}
[bread, butter] \in\ valid \implies\ valid \vdash assoc([bread,butter],[jelly])
\end{align*}
%
where the $assoc$ predicate denotes that the words must all appear in a file. 
In this way, our formalism is more expressive than the classic association rule learning problem.

The set $\keys$ is the set of unique keys from the training set (denoted \trainingSet) and the predicate $p$ is one of the classes of configuration errors.
The task of the learning algorithm is to transform a training set to a set of rules, weighted with \textit{support} and \textit{confidence}.
The set of rules learned from training set \trainingSet constitutes a specification for a configuration file to be considered correct.

The two metrics, support and confidence, are used in association rule learning, as well as other rule based machine learning techniques~\cite{han2007frequent,langley1995applications}.
We use slightly modified definitions of these, stated below, to handle arbitrary predicates as rules.
During the learning process, each rule is assigned a support and confidence to measure the amount and quality of evidence for the rule.
%
\begin{equation*}
 \textit{support(r)} = \frac{|\{C \in \trainingSet \mid S_r \cup T_r \subseteq C\}|} {|\trainingSet|}
\end{equation*}
\begin{equation*}
 \textit{confidence(r)} = \frac{|\{C \in \trainingSet \mid C\ \vdash p_r(S_r,T_r) \}| } {support(r)*|\trainingSet|}
\end{equation*}

@@

\begin{align}
Learn(\trainingSet) = \{r | \ &support(r) > t_s\ \land \nonumber \\
    & \textit{confidence(r)} > t_c\ \land \label{eq:learn}\\ 
    & S,\ T \subseteq \keys\} \nonumber
\end{align}

\subsection{Error Classes}
Each class of error forms a rule with a predicate $p$, and choices of $|S|$ and $|T|$, the sizes of the source and target keyword lists.
Ordering errors require $|S|=1, |T|=1$ and use the predicate $order$ which means if both keywords $S_0,T_0$ appear in a file, $S_0$ must come before $T_0$.
For example, the ordering error given in Sec.~\ref{sec:motiv} is expressed:
%
\begin{align*}
[\texttt{innodb\_data\_home\_dir}&] \in \validSet \implies \\
&\validSet \vdash order([\texttt{innoddb\_data\_home\_dir}],[\texttt{innodb\_data\_file\_path}])
\end{align*}
%
Missing keyword entry errors require $|S|, |T| = 1$ and use the predicate $missing$ to mean the keyword $T_0$ must appear, in any location, in the same file as the keyword $S_0$ in any configuration file.
The type rule is a set of rules over multiple predicates, where $|S|=1, |T|=1$.
These type rules take the form $S_0 \in \validSet \implies \validSet \vdash is\_([S_0],[S_0])$ where $\_$ matches all the basic types (bool, int, size, etc), as shown in Fig.~\ref{fig:allrules}.
In this case, the source and target are the same keyword, since a type constraint is only on a single keyword.

\app also supports two types of integer correlation rules, coarse-grained and fine-grained.
Both integer correlation rules are set of rules over the set of predicates $compare=\{<,=,>\}$.
Coarse grain rules require $|S|, |T| = 1$, and the predicates have the typical interpretation.
Fine-grained rules use $|S|=2,|T|=1$, and interpret the predicates such that for $k_1,\ k_2 \in S,\ k1*k2$ must have the predicate relation to $T_0$.
The integer correlation rules also use probabilistic types to prune the search space.
To avoid learning too many false positives, we restrict this rule to either $size*int=size$, or $int*int = int$, as shown in Fig.~\ref{fig:allrules}.
Without probabilistic typing, we would also learn rules which do not have a valid sematinc interpretation, for example a relation between three size keywords ($k1,k2,k3::size$) of the form $k1*k2>k3$.

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
%\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=bool]
{k_1 :: bool}
{isbool([k_1],[k_1]) :: Rule}
\and\quad
\inferrule* [Right=int]
{k_{1} :: int}
{isint([k_1],[k_1]) :: Rule}
\and\\
\inferrule* [Right=missing]
{ }
{missing([k_1],[k_2]) :: Rule}
\and\quad
\inferrule* [Right=coarse\_grain]
{k_1, k_2 :: int}
{compare([k_1],[k_2]) :: Rule}
\and\\
\inferrule* [Right=fine\_grain,before=\hspace{-1.3cm}]
{k_1,k_2,k_3 :: int}
{compare([k_1,k_2],[k_3]) :: Rule}
\and\quad
\inferrule* [Right=fine\_grain]
{k_1,k_3 :: size \\ k_2 :: int}
{compare([k_1,k_2],[k_3]) :: Rule}
\end{mathpar}
\caption{An expanded set of typing judgements for valid rules}
\label{fig:allrules}
\end{figure}
}

### 2017-04-16 OOPSLA17/secs/related.tex
fixed rule def to be entierly consistent (i hope)

\para{Association Rule Learning}
This approach is not only generalizes association rule learning, but also another popular learning strategy called \textit{sequential pattern mining}~\cite{seqMine}.
Our ordering rules are similar to the rules learned in sequential pattern mining, although can restrict our ordering rules to |S|,|T|=1 since these are the most common misconfigurations we encounter in practice.
While a major limitation to sequential pattern mining is the scalability of the problem, we escape this issue since configuration files tend to be no more than thousands of lines at a time.

>>>>

\para{Association Rule Learning}
This approach is not only generalizes association rule learning, but also another popular learning strategy called \textit{sequential pattern mining}~\cite{seqMine}.
Our ordering rules are similar to the rules learned in sequential pattern mining, although can restrict our ordering rules to $|S|,|T|=1$ since these are the most common misconfigurations we encounter in practice.
While a major limitation to sequential pattern mining is the scalability of the problem, we escape this issue since configuration files tend to be no more than thousands of lines at a time.

### 2017-04-16 OOPSLA17/secs/table2.tex
fixed rule def to be entierly consistent (i hope)

\newcommand{\tablewidth}{5.9cm}
\definecolor{Gray}{gray}{0.80}
\newcolumntype{g}{>{\columncolor{Gray}}c}

@@

\multirow{3}{*}{\parbox{\tablewidth} {\scriptsize TYPE ERROR: Expected an integer for “slow\_query\_log”}}
& \cite{type1} & 32/34 & 1/34 & 5/7   & 1/7 \\ 
& \cite{type2} & 9/20  & 2/20 & 10/11 & 2/11 \\  
& \cite{type3} & 9/19  & 2/19 & 10/11 & 2/11 \\
\hline

>>>>

\newcommand{\tablewidth}{6.5cm}
\definecolor{Gray}{gray}{0.80}
\newcolumntype{g}{>{\columncolor{Gray}}c}

@@

\multirow{3}{*}{\parbox{\tablewidth} {\scriptsize TYPE ERROR: Expected bool for ``thread\_cache\_size''}}
& \cite{type1} & 1/12  & 1/12 & 1/4   & 1/4 \\ 
& \cite{type2} & 1/9  & 1/9 & 1/1 & 1/1 \\  
& \cite{type3} & 1/8  & 1/8 & 1/3 & 1/3 \\
\hline
